## Data Science - Reproductible Research - Assignment 1

### Purpose

Write a report that answers the questions detailed below.

### Load and preprocess the data

```{r}
df_activity=read.csv("data/activity.csv")

```

### What is the mean of total number of steps taken per day?


Calculate the total number of steps taken per day.

```{r}
StepsDay=aggregate(steps ~ date, df_activity, sum, na.rm = "TRUE")

```
Make a histogram of the total number of steps taken each day.

```{r}
hist(StepsDay$steps, main="Histogram of total number of steps taken each day",
     xlab="Total Number of steps")
```
Calculate and report the mean and median of the total number of steps taken per day.

```{r}
mean(StepsDay$steps)

```
The mean of the total number of steps taken per day is 10,766.19 steps.

```{r}
median(StepsDay$steps)

```
The median of the total number of steps taken per day is 10,765 steps.

### What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r}
with(df_activity,
    plot(aggregate(steps, by=list(interval), mean, na.rm = "TRUE"),
         xlab="5 minutes interval", ylab="Average Number of steps",
         main="Average daily activity pattern", type= "l"))

```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
StepsInt=aggregate(steps ~ interval, df_activity, mean, na.rm = "TRUE")

StepsInt[which.max(StepsInt$steps),]

```
We see that this is the **835th** interval.


### Imputing missing values

There are a number of days/invervals where there are missing values.The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r}

summary(df_activity)

table(is.na(df_activity$steps))

mean(is.na(df_activity$steps))

```

We see that the missing values are present in the 'step' variable only and that they represent **13 percent** of the total values.

```{r}
sum(is.na(df_activity$steps))
```

The total number of missing values in the dataset is **2304**.

#### How are these NAs distributed?

```{r}
data=aggregate(is.na(df_activity$steps) ~ date, df_activity, sum)

colnames(data)=c("date", "sumNA")

data[data$sumNA > 0,]

```
We notice that all intervals for the full day (24*60/5=288) are missing for 8 dates.

#### Strategy for filling in all the missing values.

The strategy that I have chosen is to use the mean for that particular missing 5-minute interval.

Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r}

df_activity2=df_activity

df_activity2$steps = ifelse(is.na(df_activity2$steps),StepsInt$steps[match(df_activity2$interval, StepsInt$interval)],df_activity2$steps)


```

Make a histogram of the total number of steps taken each day.

```{r}
StepsDay2=aggregate(steps ~ date, df_activity2, sum, na.rm = "TRUE")

hist(StepsDay2$steps, main="Histogram of total number of steps taken each day",
     xlab="Total Number of steps")

```

Calculate and report the mean and median of the total number of steps taken per day.

```{r}
mean(StepsDay2$steps)
median(StepsDay2$steps)

```

####  Do these values differ from the estimates from the first part of the assignment?

The mean and the median are equal to **10,766.19 steps***. 

We notice that our imputation strategy slightly impacted the median, although the mean remained equal.

#### What is the impact of imputing missing data on the estimates of the total daily number of steps?

We would have expected that the imputation would have introduced a bias.

In the chosen strategy the bias is negligible for the given dataset.

### Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}

MyLocale <- Sys.getlocale("LC_TIME")

Sys.setlocale("LC_TIME", "English")

IsWeekDay=function(date){
    givenday=weekdays(date)
    if(givenday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) 
        return("weekday")
    else if (givenday %in% c("Saturday", "Sunday"))
        return("weekend")
}
df_activity2$daytype=sapply(as.Date(df_activity2$date), IsWeekDay)

```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
AvSteps=aggregate(steps ~ interval + daytype, df_activity2, mean)

library(ggplot2)

ggplot(AvSteps, aes(interval, steps)) + geom_line() + facet_grid(daytype ~ .) + 
    xlab("5 minutes interval") + ylab("Average Number of steps")


Sys.setlocale("LC_TIME", MyLocale)

```


